---
title: "Team Charlotte Weekly Report"
author: "Adi & David"
format: gfm
editor: visual
---

## Research Question

What is the impact of particulate matter (PM2.5) and other environmental factors on human health for individuals living in proximity to the Blue Line light rail in Charlotte, North Carolina.

## Hypothesis

We hypothesize that:

1.  The concentration of PM2.5 is higher near the Blue Line light rail stations compared to areas further away.
2.  Proximity to major traffic intersections, industrial areas, power plants, and airports exacerbates the level of PM

## Context

-   Area of study: Blue Lynx Light Rail, Charlotte, North Caroline

    -   Table with corresponding above stations

        ```{r}
        library("knitr")

        stations_table <- read.csv("geocoded_data.csv")
        kable(stations_table)
        ```

-   Time Frame: November 2003 - November 2011

-   Factors associated with PM2.5

    -   Airport

    -   Power Plant(s)

    -   Factories

    -   Major Intersection

        -   Table with corresponding above factors

            ```{r}
            factors_table <- read.csv("new_pm_geocoded_data2.csv")
            kable(factors_table)
            ```

### Team Tasks

-   Chose a City for the Research

    -   Charlotte, North Carolina

-   Created a [spreadsheet](https://docs.google.com/spreadsheets/d/1ZFwtPYZDcX4K_ywbQWh51AtqWkiJgJTiqacHCCI4Mdc/edit#gid=0) that tracks the following:

    1.  Station Name

    2.  Address

    3.  Built Date

    4.  Coordinates

    5.  Location

    6.  Parking Availability

-   Completed GeoCoding to find coordinates of all the stations in Lynx Blue Line, Charlotte, NC

-   Created an account, project and use the Google GeoCoding API

-   Converted coordinates into spatial objects and plot them

-   Created buffers around stations and calculate daily PM2.5

-   Found sources around stations that contribute towards the PM2.5 levels

-   Created an Interactive Map with buffer zones around stations \[optional\]

### Individual Tasks

-   Complete Tidyverse DataCamp: David Guan

-   Complete Spatial Training: Adityaraj Padmanabhan

## 
